---
title: "02_PairwiseCompPart2"
output: html_document
---

### Carlvalho et al., 

***
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

#### Setup and load data

```{r load libraries, include = FALSE}
library(tidyverse)
require(tidyverse)
library(here)
library(flextable)
library(jtools)
library(kableExtra)
```


```{r load data, include = FALSE}
here()
df <- read.csv(here("..","..","01_data","all_GPLxBrazil.csv"), header = T, sep = ",")
names(df)
#glimpse(df)

```




## How many individuals in each treatment were exposed but uninfected?

```{r data overview , include = FALSE}

infected <- df %>% 
  group_by(Treatment, StrainID, Lineage_type, FieldN_2, Infected) %>%
  dplyr::filter(FieldN_2 == "S1") %>% 
  mutate(n_obs = n()) %>% # count the number of observations per group
  mutate(bd1 = mean(TotalBd, na.rm = TRUE)) %>% 
  mutate(StrainID = as.factor(StrainID)) %>% 
  ungroup() %>% 
  dplyr::select(Treatment, StrainID, Lineage_type, bd1, n_obs) %>%
  distinct() %>% as.data.frame()
infected
```




```{r early establishment single-genotype infections, include = FALSE}

# First, look at single-genotype infections, so subset
# Examine early establishment - that is, the Bd load in Swab 1.

bds1 <- df %>% 
  dplyr::filter(Treatment!= "Coinfected") %>% 
  group_by(Treatment, ID, StrainID, Lineage_type, FieldN_2) %>%
  dplyr::filter(FieldN_2 == "S1") %>% 
  mutate(n_obs = n()) %>% # count the number of observations per group
  mutate(bd1 = mean(TotalBd, na.rm = TRUE)) %>% 
  mutate(StrainID = as.factor(StrainID)) %>% 
  ungroup() %>% 
  dplyr::select(Treatment, ID, StrainID, Lineage_type, bd1) %>%
  distinct() %>% as.data.frame()
bds1

with(bds1, hist(log(bd1)))
with(bds1, hist(bd1))

```

```{r check the ditribution, include = FALSE}
require(fitdistrplus) # 
# 
fit.norm     <- fitdist(bds1$bd1,"norm", method = c("mle"))  
plot(fit.norm)
fit.gamma    <- fitdist(bds1$bd1,"gamma", method = c("mle"))#best fit
plot(fit.gamma)
fit.weibull  <- fitdist(bds1$bd1,"weibull", method = c("mle"))
plot(fit.weibull)
fit.lnorm    <- fitdist(bds1$bd1,"lnorm", method = c("mle"))
plot(fit.lnorm)

```


```{r run the model, include = TRUE}

m1 <- glm(bd1 ~ StrainID, family = Gamma(link = "log"), data = bds1)
summ(m1)
#summ.lm(m1)

```


## update - clarify why this contrast doesn't work based on Crawley and Bolker
```{r trial contrast, include = FALSE}
with(bds1, levels(StrainID))
coef(m1)


contrasts(bds1$StrainID) <- cbind(c(-1, -1, -1, 4, -1),
                                  c(0,0,0,0,0),
                                  c(0,0,0,0,0),
                                  c(0,0,0,0,0))
contrasts(bds1$StrainID)
```

```{r trial conrast summary, include=TRUE}
m1 <- glm(bd1 ~ StrainID, family = Gamma(link = "log"), data = bds1)
summ(m1)
plot(m1)

```


```{r subset 2: compare coinfections with P2 to P2 single-genotype infections,  include = FALSE}
# Second, compare coinfections with P2 to P2 single-genotype infections
## Does coinfection change competitiveness (early establishment of P1)?
bds2 <- df %>% 
  group_by(Treatment, Infected, ID, StrainID, Lineage_type, FieldN_2) %>%
  dplyr::filter(StrainID == "P2") %>%
  dplyr::filter(FieldN_2 == "S1") %>% 
  mutate(n_obs = n()) %>% # count the number of observations per group
  mutate(bd1 = mean(TotalBd, na.rm = TRUE)) %>% 
  mutate(StrainID <- factor(StrainID)) %>% 
  dplyr::select(FieldN_2, Infected, Treatment, ID, StrainID, Lineage_type, bd1) %>%
  distinct() %>% as.data.frame()
bds2



bds3 <- df %>% 
  group_by(Treatment, Infected, ID, StrainID, Lineage_type, FieldN_2) %>%
  dplyr::filter(Treatment ==  "Coinfected") %>%
  dplyr::filter(StrainID != "E2-P1") %>% 
  dplyr::filter(StrainID != "E1-P1") %>% 
  dplyr::filter(StrainID != "E1-P1-R") %>% 
  dplyr::filter(StrainID != "E1-R") %>% 
  dplyr::filter(StrainID != "E2-P1-R") %>% 
  dplyr::filter(StrainID != "E2-R") %>% 
  dplyr::filter(StrainID != "P1-R") %>% 
  dplyr::filter(FieldN_2 == "S1") %>% 
  mutate(n_obs = n()) %>% # count the number of observations per group
  mutate(bd1 = mean(TotalBd, na.rm = TRUE)) %>% 
  mutate(StrainID <- factor(StrainID)) %>% 
  dplyr::select(FieldN_2, Infected, Treatment, ID, StrainID, Lineage_type, bd1) %>%
  distinct() %>% as.data.frame()
bds3


```


```{r merge datasets for pairwise contrasts 1, include = FALSE}
 
bds5 <- bind_rows(bds2, bds3) %>% 
  drop_na()
bds5

```


```{r check the distribution for data subset for coinfections, include = FALSE}
# require(fitdistrplus) 
# 
# with(bds5, factor(StrainID))
# 
# with(bds5, hist(log(bd1)))
# with(bds5, hist(bd1))
# 
# fit.norm     <- fitdist(bds5$bd1,"norm", method = c("mle"))
# plot(fit.norm)
# #fit.gamma    <- fitdist(bds5$bd1,"gamma", method = c("mle"))#best fit
# #plot(fit.gamma)
# fit.weibull  <- fitdist(bds5$bd1,"weibull", method = c("mle"))
# plot(fit.weibull)
# fit.lnorm    <- fitdist(bds5$bd1,"lnorm", method = c("mle"))
# plot(fit.lnorm)


```


```{r run the candidate models for coinfections, include = TRUE}


m1 <- glm(bd1 ~ Treatment, data = bds5)
summ(m1)
#summary.lm(m1)
```


## can't use this for this subset since there are zeros in the P2 subset
* since these zeros are informative, we will keep m1 even though it's a bit different from the P1 comparison (which did not have zeros)

```{r, include = FALSE}
# m2 <- glm(log(bd1) ~ Treatment,  data = bds4)
# summary(m2)
# summary.lm(m2)

# AIC(m1, m2)
```


## No difference here, so we will go with the fits above...
* above, we see that the log normal fits better - esp. if we consider the Q-Q plot. 


```{r set up the next contrasts, include = FALSE}

# bds5$StrainID <- as.factor(bds5$StrainID)
# 
# m1 <- glm(bd1 ~ as.factor(StrainID), data = bds5)
# summary.lm(m1)
# 
# with(bds5, levels(StrainID))
# coef(m1)
# 
# contrasts(bds5$StrainID) <- cbind(c(0, 0, 0, 0, 0, 0),
#                                   c(1, 0, 0, 0, -1, 0),
#                                   c(0, 0, 0, 0, -1, 1),
#                                   c(0, 0, 0, 0, 0, 0),
#                                   c(0, 0, 0, 0, 0, 0),
#                                   c(0, 0, 0, 0, 0, 0))
# contrasts(bds5$StrainID)
# 
# m1 <- glm(bd1 ~ StrainID,  data = bds5)
# summary.lm(m1)
```

## Fourth, transmission potential (total Bd load)
* Since for Total Bd Load, we are examining whether co-infection had an effect....in that case, we are actually only interested in animals that became infected. 
  + Remove animals that were exposed but uninfected 

```{r transmission potential (total Bd load), include = FALSE}

trans <- df %>%
  group_by(ID, Strains, StrainID, Treatment, Lineage_type, Lineage) %>%
  dplyr::filter(TotalBd > 0) %>%
  mutate(meanbd  = mean(TotalBd, na.rm = TRUE),
         logbd   = mean(log(TotalBd), na.rm = TRUE),
         meanGPL = mean(GPL_pg, na.rm = TRUE),
         totGPL  = sum(GPL_pg, na.rm = TRUE),
         meanBZL = mean(Brazil_pg, na.rm = TRUE),
         totBZL  = sum(Brazil_pg, na.rm = TRUE)) %>%
  mutate(dod2  = sum(Day_swab, na.rm = T)) %>%
  dplyr::select(ID,  StrainID, Treatment, Lineage_type, Lineage, meanbd, logbd, meanGPL, totGPL, meanBZL, totBZL, dod2) %>%
  distinct() %>% as.data.frame()
head(trans)



trans1 <- df %>%
  group_by(ID, Strains, StrainID, Treatment, Lineage_type, Lineage) %>%
  dplyr::filter(TotalBd > 0) %>%
  dplyr::filter(StrainID == "P2") %>%
  mutate(meanbd  = mean(TotalBd, na.rm = TRUE),
         logbd   = mean(log(TotalBd), na.rm = TRUE),
         meanGPL = mean(GPL_pg, na.rm = TRUE),
         totGPL  = sum(GPL_pg, na.rm = TRUE),
         meanBZL = mean(Brazil_pg, na.rm = TRUE),
         totBZL  = sum(Brazil_pg, na.rm = TRUE)) %>%
  mutate(dod2  = sum(Day_swab, na.rm = T)) %>%
  dplyr::select(ID,  StrainID, Treatment, Lineage_type, Lineage, meanbd, logbd, meanGPL, totGPL, meanBZL, totBZL, dod2) %>%
  distinct() %>% as.data.frame()
head(trans1)


trans2 <- df %>%
  group_by(ID, Strains, StrainID,Treatment, Lineage_type, Lineage) %>%
  dplyr::filter(TotalBd > 0) %>%
  dplyr::filter(Treatment ==  "Coinfected") %>%
  dplyr::filter(StrainID != "E2-P1") %>%
  dplyr::filter(StrainID != "E1-P1") %>%
  dplyr::filter(StrainID != "E1-P1-R") %>%
  dplyr::filter(StrainID != "E1-R") %>%
  dplyr::filter(StrainID != "E2-p1-R") %>%
  dplyr::filter(StrainID != "E2-R") %>%
  dplyr::filter(StrainID != "P1-R") %>%
  mutate(meanbd  = mean(TotalBd, na.rm = TRUE),
         logbd   = mean(log(TotalBd), na.rm = TRUE),
         meanGPL = mean(GPL_pg, na.rm = TRUE),
         totGPL  = sum(GPL_pg, na.rm = TRUE),
         meanBZL = mean(Brazil_pg, na.rm = TRUE),
         totBZL  = sum(Brazil_pg, na.rm = TRUE)) %>%
  mutate(dod2  = sum(Day_swab, na.rm = T)) %>%
  dplyr::select(ID, Strains, StrainID, Treatment, Lineage_type, Lineage, meanbd, logbd, meanGPL, totGPL, meanBZL, totBZL, dod2) %>%
  distinct() %>% as.data.frame()
head(trans2)

```


```{r merge datasets for pairwise contrasts 2, include = FALSE}
trans3 <- bind_rows(trans1, trans2)
trans3
#End
```



```{r check distribiutions for virulence on lifespan, include = FALSE}
# fairly normal distribution
with(trans3, hist(log(meanbd)))
with(trans3, hist(meanbd))

# check the distribution
require(fitdistrplus) # 
# 
fit.norm     <- fitdist(trans3$meanbd,"norm", method = c("mle"))  
plot(fit.norm)
fit.gamma    <- fitdist(trans3$meanbd,"gamma", method = c("mle"))#best fit
plot(fit.gamma)
fit.weibull  <- fitdist(trans3$meanbd,"weibull", method = c("mle"))
plot(fit.weibull)
fit.lnorm    <- fitdist(trans3$meanbd,"lnorm", method = c("mle"))
plot(fit.lnorm)



```


```{r candidate models, include = TRUE}
m1 <- glm(meanbd ~ Treatment, family = Gamma(link = "log"), data = trans3)
summ(m1)
#summary.lm(m1)
```

## Now, Virulence on lifespan 

```{r subset data for virulence on lifespan, include = FALSE}
# Third, compare coinfections with P1 to P1 single-genotype infections
## Does coinfection change transmission potential



lifespan1 <- df %>%
  group_by(ID, Infected, Treatment, StrainID, Lineage_type, Lineage) %>%
  dplyr::filter(FieldN_2 == "S2") %>% # subset for final, otherwise, we'll double count
  dplyr::filter(Infected == "1") %>% 
  mutate(dod2  = sum(Day_swab, na.rm = T)) %>%
  dplyr::select(ID, Treatment, StrainID, Lineage_type, Lineage, dod2) %>%
  distinct() %>% as.data.frame()
lifespan1



lifespan2 <- lifespan1 %>% 
  group_by(ID, Infected, Treatment, StrainID, Lineage_type) %>%
  mutate(meandod = mean(dod2, na.rm = TRUE)) %>% 
  dplyr::select(ID, Infected, Treatment, StrainID, Lineage_type, meandod) %>%
  distinct() %>% as.data.frame()
lifespan2




lifespanP2 <- lifespan1 %>% 
  group_by(ID, Infected, Treatment, StrainID, Lineage_type) %>%
  dplyr::filter(StrainID == "P2") %>% 
  mutate(meandod = mean(dod2, na.rm = TRUE)) %>% 
  dplyr::select(ID, Infected, Treatment, StrainID, Lineage_type, meandod) %>%
  distinct() %>% as.data.frame()
lifespanP2


lifespanP2Co <- lifespan1 %>% 
  group_by(ID, Infected, Treatment, StrainID, Lineage_type) %>%
  dplyr::filter(Treatment ==  "Coinfected") %>%
  dplyr::filter(StrainID != "E2-P1") %>% 
  dplyr::filter(StrainID != "E1-P1") %>% 
  dplyr::filter(StrainID != "E1-P1-R") %>% 
  dplyr::filter(StrainID != "E1-R") %>% 
  dplyr::filter(StrainID != "E2-P1-R") %>% 
  dplyr::filter(StrainID != "E2-R") %>% 
  dplyr::filter(StrainID != "P1-R") %>% 
  mutate(meandod = mean(dod2, na.rm = TRUE)) %>% 
  dplyr::filter(meandod  !=  "0") %>%
  dplyr::select(ID, Infected, Treatment, StrainID, Lineage_type, meandod) %>%
  distinct() %>% as.data.frame()
lifespanP2Co

```


```{r merge datasets for pairwise contrasts 3, include = FALSE}

P2life <- bind_rows(lifespanP2, lifespanP2Co)
P2life

```



```{r check distributions, include = FALSE}
# fairly normal distribution
with(P2life, hist(meandod))
with(P2life, hist(log(meandod)))
with(P2life, hist(sqrt(meandod)))


with(P2life, range(meandod))

library(fitdistrplus)
require(fitdistrplus) # 
# 
fit.norm     <- fitdist(P2life$meandod,"norm", method = c("mle"))  
plot(fit.norm)
fit.gamma    <- fitdist(P2life$meandod,"gamma", method = c("mle"))#best fit
plot(fit.gamma)
fit.weibull  <- fitdist(P2life$meandod,"weibull", method = c("mle"))
plot(fit.weibull)
fit.lnorm    <- fitdist(P2life$meandod,"lnorm", method = c("mle"))
plot(fit.lnorm)


```


```{r run the candidate models, include = FALSE}

lm1<- glm(meandod ~ Treatment, data = P2life)
lm1
#summary.lm(lm1)
summ(lm1)
plot(lm1)

# no evidence of hetero
library(lmtest)
lmtest::bptest(lm1) # reject null = hetero

m5 <- glm(meandod  ~ Treatment, data = P2life)
m6 <- glm(log(meandod)  ~ Treatment, data = P2life)


AIC(m5, m6) 

summ(m6)
#summary.lm(m6)
plot(m6)
m5

```


```{r summary tables for output, include=TRUE}
summ(lm1)
plot(lm1)
summ(m6)
plot(m6)
```





```{r data summaries for inspection, include = FALSE}
relP2 <- df %>%
  group_by(ID, StrainID, Lineage_type, Lineage, Treatment, FieldN_2,  Day_swab, TotalBd) %>%
  filter(Treatment == "Coinfected") %>% 
  mutate(meanGPL = mean(GPL_pg, na.rm = T)) %>%
  mutate(totGPL  = sum(GPL_pg, na.rm = T)) %>%
  mutate(meanBZL = mean(Brazil_pg, na.rm = T)) %>%
  mutate(totBZL  = sum(Brazil_pg, na.rm = T)) %>%
  mutate(meanH   = mean(Hybrid , na.rm = T)) %>%
  mutate(totH    = sum(Hybrid , na.rm = T)) %>%
  dplyr::select(ID, StrainID, Lineage_type, Lineage, Treatment, FieldN_2,  Day_swab, meanGPL, totGPL, meanBZL, totBZL, meanH, totH, TotalBd) %>%
  distinct() %>% as.data.frame()
head(relP2)


relP2a <- relP2 %>% 
  group_by(ID, StrainID, Lineage_type, Lineage, Treatment, FieldN_2, Day_swab) %>% 
  mutate(totbd   = sum(totGPL + totBZL + totH, na.rm = T)) %>%
  mutate(relgpl  = (totGPL/TotalBd)) %>%
  group_by(StrainID) %>%
  mutate(n_obs = n()) %>% # count the number of observations per group
  dplyr::select(ID, StrainID, Lineage_type, Lineage, Treatment, FieldN_2, Day_swab, relgpl) %>%
  distinct() %>% as.data.frame()
head(relP2a)


relP2b<- relP2a %>% 
  group_by(StrainID, Treatment, FieldN_2) %>%
  filter(Treatment == "Coinfected") %>% 
  filter(FieldN_2 == "S2") %>% 
  mutate(dayofdeath = mean(Day_swab, na.rm = TRUE)) %>% 
  mutate(meanrelgpl = mean(relgpl, na.rm = TRUE)) %>% 
  mutate(nobs = n()) %>% 
  mutate(relgplse = sd(relgpl, na.rm = TRUE)/sqrt(nobs)) %>%
  mutate(meandod = mean(Day_swab, na.rm = TRUE)) %>% 
  mutate(nobs = n()) %>% 
  mutate(dodse = sd(Day_swab, na.rm = TRUE)/sqrt(nobs)) %>%
  dplyr::select(StrainID, Lineage_type, meandod, dodse, meanrelgpl, relgplse, dayofdeath) %>%
  distinct() %>% as.data.frame()
relP2b


```

