---
title: "03_MortalityStats"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## Carlvalho et al., 
***

### Setup and load data

```{r load libraries, echo = FALSE, include = FALSE}
library(tidyverse)
require(tidyverse)
library(jtools)
library(multcomp)
library(mgcv)
library(here)
```



```{r load data, echo = FALSE, include = FALSE}
here()
df <- read.csv(here("all_GPLxBrazil.csv"), header = T, sep = ",")
names(df)
glimpse(df)
names(df)
glimpse(df)

```

## Begin Figure 7: % Mortality - virulence
* [calculate % Mortality](https://campus.datacamp.com/courses/basic-statistics/lab-4-probability-distributions?ex=10)
  + the mean is calculated by multiplying the number of trials, n, by the prob. of a success, p. 
    - the prob of success (success or failure), p, is $2/n$
* sd is the following: $\sqrt(n*p*(1-p))$
* mean = n * p
* calculate the total number of trials, n: total animals/strain
* `mean_chance <- 2/n`
* [calculate binomial mean and stdev](https://mse.redwoods.edu/darnold/math15/spring2013/R/Activities/MeanVarStd.html)

```{r data wrangling , echo = FALSE, include = FALSE}
dead <- df %>% 
  group_by(StrainID, Lineage_type,  FieldN_2) %>%
  # count only second swab otherwise, the sample size is inflated! 
  dplyr::filter(FieldN_2 == "S2") %>%  
  mutate(totInf  = sum(Infected, na.rm = T),
            totSusc = sum(Susceptible, na.rm = T),
            N = sum(totSusc + totInf),
            totDead = sum(Dead, na.rm = TRUE),
            probdie =  totDead /(totInf + totSusc)) %>% 
  mutate(meandead = sum(1*probdie)) %>% 
  mutate(variance = sum((1^2)*meandead) - (meandead^2)) %>% 
  mutate(stdevD = sqrt(variance)) %>% 
  dplyr::select(StrainID, Lineage_type, totDead, totSusc, totInf) %>%
  distinct() %>% as.data.frame()
dead


#cbind (dead, Susceptible)

```




```{r try GAMS, echo = FALSE, include = TRUE}
# see: https://m-clark.github.io/generalized-additive-models/case_for_gam.html

y <- with(dead, cbind(totDead, (totSusc + totInf)))

# run the full model
m6 <- gam(y ~ StrainID,  data = dead, family = binomial)
# m1 diagnostics
summary(m6)
gam.check(m6, k.rep = 1000)



```

## Contrast Matrix
* use -1 for the baseline group (here, comparing load in coninfection to the the most virulent strain in the single infections, based on host lifespan). 
  + This says, does coinfection decrease virulence relative to single infections?

```{r define and analyze contrast matrix, echo = FALSE, include = FALSE}


with(dead, levels(StrainID))




# USE Compare R single infection (reference strain) against each other single strain infection 
cmat <- rbind("P1.vs.E1-P1" = c1 <- c(0,1,0,0,0,0,0,0,-1,0,0,0,0),
              "E2-P1.vs.E2" = c2 <- c(0,0,0,0,-1,1,0,0,0,0,0,0,0),
              "E1-P2.vs.P2"  = c3 <- c(0,0,0,0,0,1,0,0,0,0,-1,0,0),
              "P2-E2.vs.E2" = c4 <- c(0,0,0,0,-1,0,1,0,0,0,0,0,0),
              "E1.vs.E2" = c5 <- c(1,0,0,0,-1,0,0,0,0,0,0,0,0),
              "E2.vs.P1" = c6 <- c(0,0,0,0,-1,0,0,0,1,0,0,0,0),
              "P1.vs.P2" = c7 <- c(0,0,0,0,0,0,0,0,-1,0,1,0,0),
              "R.vs.E1." = c8 <- c(1,0,0,0,0,0,0,0,0,0,0,0,-1),
              "R.vs.E2." = c9 <- c(0,0,0,0,-1,0,0,0,0,0,0,0,-1),
              "R.vs.P1." = c10 <- c(0,0,0,0,0,0,0,0,1,0,0,0,-1),
              "R.vs.P2." = c11 <- c(0,0,0,0,0,0,0,0,0,0,1,0,-1),
              "R-E2.vs.E2" = c12 <- c(0,0,0,0,-1,0,0,1,0,0,0,0,0),
              "R-E1.vs.E1" = c13 <- c(-1,0,0,1,0,0,0,0,0,0,0,0,0),
              "R-P1.vs.P1" = c14 <- c(0,0,0,0,0,0,0,0,-1,1,0,0,0),
              "R-P2.vs.P2" = c1 <- c(0,0,0,0,0,0,0,0,0,0,0,1,-1))
cmat


#summary(glht(m4, linfct = cmat))



```

