---
title: "04_Plots"
output: html_document
date: "2022-09-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Carlvalho et al., 

***

# Setup and load data 
```{r load libraries, include = FALSE}

library(tidyverse)
require(tidyverse)
library(cowplot)  
library(extrafont)
library(patchwork)
library(here)
here()

```


```{r load data, include = FALSE}

df <- read.csv(here("all_GPLxBrazil.csv"), header = T, sep = ",")
names(df)
glimpse(df)

```



```{r figure parameters, include = FALSE}

shape3 <- c(21, 22, 23, 24, 25, 21, 21, 21, 21, 21, 22, 22, 22, 22, 22, 23, 24)
shapeP1 <- rep(21, 6)
shapeP2 <- rep(22, 6)

col3 <- c("#009E73", "#785EF0", "#FE6100","#FFB000","#DC267F",  "continuous")
col3a <- c("white", "white", "white","white","white")

colP1 <- c("#009E73", "#FE6100", "#FE6100", "#FFB000", "#FFB000","#DC267F")
colP1a <- c("white", "#FE6100", "#FE6100", "#FFB000", "#FFB000","#DC267F")

colP2 <- c("#785EF0", "#FE6100", "#FE6100", "#FFB000", "#FFB000", "#DC267F")
colP2a <- c("white", "#FE6100", "#FE6100", "#FFB000", "#FFB000", "#DC267F")


```


## Figure 1 - Proportion infected

```{r prep data for FIGURE 1 - Proportion infected, include = FALSE}
# calculate binomial mean and stdev
# https://mse.redwoods.edu/darnold/math15/spring2013/R/Activities/MeanVarStd.html

#binom.test(totInf, totalN, 0.5, alternative = "two.sided", conf.level = 0.95)
inf <- df %>% 
  dplyr::group_by(StrainID, Treatment, Lineage_type, Lineage, FieldN_2) %>%
  # count only second swab otherwise, the sample size is inflated! 
  dplyr::filter(FieldN_2 == "S2") %>%  
  mutate(nobs = n()) %>% 
  summarize(totInf  = sum(Infected, na.rm = T),
            totSusc = sum(Susceptible, na.rm = T),
            freq.inf  = totInf/(totInf + totSusc)) %>% 
  mutate(meaninf = sum(1*freq.inf)) %>% 
  mutate(variance = sum((1^2)*meaninf) - (meaninf^2)) %>% 
  mutate(stdev = sqrt(variance)) %>% 
  dplyr::select(StrainID, Treatment, Lineage_type, Lineage, totInf, totSusc, freq.inf, meaninf, stdev) %>%
  distinct() %>% as.data.frame()
inf

pinfP1 <- as.list(subset(inf, StrainID == "P1", meaninf))
pinfP1

pinfP2 <- as.list(subset(inf, StrainID == "P2", meaninf))
pinfP2

```


```{r figure 1 initialization, include = FALSE}

# Re-order Lineage type for the correct figure order
inf$Lineage_type <- factor(inf$Lineage_type, c("Enzootic","Panzootic","Recombinant","E-P Coinfection","R Coinfection"))



inf <- inf %>%
  mutate(StrainID = fct_relevel(StrainID, 
                                "P1", "P2", "E1","E2","R",
                                "P1-E1","P1-E1-R","P1-E2","P1-E2-R","P1-R",
                                "P2-E1", "P2-E1-R","P2-E2","P2-E2-R","P2-R"))


```

### Figure 1 part 1
```{r figure 1.1, include = TRUE}
p2 <- inf %>%
  #dplyr::filter(Grouping == "1") %>%
  group_by(Lineage_type, StrainID) %>%
  ggplot(aes(x = StrainID, y = freq.inf, color = as.factor(StrainID), fill = as.factor(StrainID), group = StrainID)) +
  geom_hline(yintercept =  as.numeric(pinfP1), linetype = "dashed", color = "#009E73", size = 0.5) +
  geom_hline(yintercept = as.numeric(pinfP2), linetype = "dotted", color = "#785EF0", size = 0.5) +
  geom_point(size = 2.50, aes(shape = StrainID, color = StrainID)) +
  geom_errorbar(aes(ymin = freq.inf - stdev, ymax = freq.inf + stdev), width = 0,  size = 0.10) +
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +
  theme(plot.margin = unit(c(0,0,0,0), "cm")) +
  geom_line(alpha = 0.45) +
  ## Need to add more color values to col3 and col3a, then uncomment
  ## for now just using default R colors
  # scale_color_manual(values = col3, name = "StrainID") +
  # scale_fill_manual(values = col3a, name = "StrainID") +
  scale_shape_manual(values = shape3) +
  ylim(0, 1.60) +
  theme_bw() +
  theme(legend.position = 'none', panel.grid.major = element_blank(),
        panel.grid.minor =  element_blank()) +
  xlab("") + ylab("% Infected") +
  theme(axis.title = element_text(size = 10)) +
  theme(text = element_text(size = 10)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5))+
  #theme(axis.text.x =  element_blank()) +
  theme(text = element_text(size = 10)) +
  theme(axis.ticks = element_line(size = 0.10),
        axis.ticks.length = unit(0.10, "cm"))
p2 + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
```


### Fig 1 part 2
```{r figure 1.2,  echo = FALSE, include = TRUE}
p3 <- inf %>%
  #dplyr::filter(PlotGroup == "1") %>%
  group_by(Lineage_type, StrainID) %>%
  ggplot(aes(x = StrainID, y = freq.inf, color = as.factor(StrainID), fill = as.factor(StrainID), group = StrainID)) +
  geom_hline(yintercept =  as.numeric(pinfP1), linetype = "dashed", color = "#009E73", size = 0.5) +
  geom_point(size = 2.50, aes(shape = StrainID, color = StrainID)) +
  geom_errorbar(aes(ymin = freq.inf - stdev, ymax = freq.inf + stdev), width = 0,  size = 0.10) +
  geom_rect(xmin = 0.85, xmax = 1.15, ymin = -Inf, ymax = Inf,
            fill = "darkgrey", color = "NA", alpha = 0.05)+
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +
  theme(plot.margin = unit(c(0,0,0,0), "cm")) +
  geom_line(alpha = 0.45) +
  ## Need to add more color IDs to colP1 and colP1a, as well as more shapes to shapeP1
  ## for now using default R colors and shape3 
  #scale_color_manual(values = colP1, name = "StrainID") +
  #scale_fill_manual(values = colP1a, name = "StrainID") +
  #scale_shape_manual(values = shapeP1) +
  scale_shape_manual(values = shape3)+
  ylim(0, 1.60) +
  theme_bw() +
  theme(legend.position = 'none', panel.grid.major = element_blank(),
        panel.grid.minor =  element_blank()) +
  xlab("") + ylab("") +
  theme(axis.title = element_text(size = 10)) +
  theme(text = element_text(size = 10)) +
  theme(axis.text.y =  element_blank()) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5))+
  #theme(axis.text.x =  element_blank()) +
  theme(axis.ticks = element_line(size = 0.10) ,
        axis.ticks.length = unit(0.10, "cm"))
p3
```

### Fig 1 part 3
```{r Figure 1.3,  echo = FALSE, include = TRUE}
p4 <- inf %>% 
  #dplyr::filter(PlotGroup == "2") %>% 
  group_by(Lineage_type, StrainID) %>%
  ggplot(aes(x = StrainID, y = freq.inf, color = as.factor(StrainID), fill = as.factor(StrainID), group = StrainID)) +
  geom_hline(yintercept = as.numeric(pinfP2), linetype = "dotted", color = "#785EF0", size = 0.5) +
  geom_point(size = 2.50, aes(shape = StrainID, color = StrainID)) +
  geom_errorbar(aes(ymin = freq.inf - stdev, ymax = freq.inf + stdev), width = 0,  size = 0.10) +
  geom_rect(xmin = 0.85, xmax = 1.15, ymin = -Inf, ymax = Inf, 
            fill = "darkgrey", color = "NA", alpha = 0.05)+
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +
  theme(plot.margin = unit(c(0,0,0,0), "cm")) +
  geom_line(alpha = 0.45) +
  ## Same issue as previous chunk
  #scale_color_manual(values = colP2, name = "StrainID") +
  #scale_fill_manual(values = colP2a, name = "StrainID") +
  #scale_shape_manual(values = shapeP2) + 
  scale_shape_manual(values = shape3) + 
  ylim(0, 1.60) +
  theme_bw() +
  theme(legend.position = 'none', 
        panel.grid.major = element_blank(), 
        panel.grid.minor =  element_blank()) +
  xlab("") + ylab("") +
  theme(axis.title = element_text(size = 10)) +
  theme(text = element_text(size = 10)) +
  theme(axis.text.y =  element_blank()) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5))+
  #theme(axis.text.x =  element_blank()) +
  theme(axis.ticks = element_line(size = 0.10) , 
        axis.ticks.length = unit(0.10, "cm"))
p4
```

### Final Fig 1

```{r combine figures for FIGURE 1, include = TRUE}
 


# Put the plots together
BdPlot1a <- cowplot::plot_grid(p2, p3, p4, 
            nrow = 1, align = "hv", 
            labels = c("A", "B", "C"),
            label_size = 10)
BdPlot1a



BdPrev <- p2 + p3 + p4 +
  plot_layout(ncol = 3)


# BdPrev + plot_annotation(tag_levels = c('A', '1'), tag_prefix = '', tag_sep = '', 
#          tag_suffix = '') & 
#   theme(plot.tag.position = c(0.12, 0.98),
#         plot.tag = element_text(size = 12, hjust = 0, vjust = 0))
#   
ggsave("BdPrev.png", height = 3.5, width = 10, dpi = 600)

plot1 <- tiff("BdPrev.tiff", width = 15, height = 22, units = "cm", res = 300, pointsize = 10) 
plot1
dev.off()  

BdPrev

```

## FIGURE 2A - Potential  Bd load Swab 1 - how quickly does each strain establish?


```{r FIGURE 2A initialization, include = FALSE}


#### FIGURE 2 part A - Potential  Bd load Swab 1 - how quickly does each strain establish? ####
bds1 <- df %>% 
  group_by(StrainID, Lineage_type, FieldN_2) %>%
  filter(FieldN_2 == "S1") %>% 
  mutate(n_obs = n()) %>% # count the number of observations per group
  mutate(bd1 = mean(TotalBd, na.rm = TRUE)) %>% 
  mutate(bd1se = sd(TotalBd, na.rm = TRUE)/sqrt(n())) %>%
  dplyr::select(StrainID, Lineage_type, bd1, bd1se) %>%
  distinct() %>% as.data.frame()
bds1

bds1P1 <- as.list(subset(bds1, StrainID == "P1", bd1))
bds1P1

bds1P2 <- as.list(subset(bds1, StrainID == "P2", bd1))
bds1P2

# now, plot bd load at swab 1
# Re-order Lineage type for the correct figure order
bds1$Lineage_type <- factor(bds1$Lineage_type, c("Enzootic","Panzootic","Recombinant","E-P Coinfection","R Coinfection"))
genolabel <- c("Single Genotype Infections", "Enzootic-Panzootic Coinfections", "Recombinant-E or -P Coinfections", "Tripple Genotype (E-P-R) Coinfections") 
names(genolabel) <- c("1", "2", "3", "4")



bds1 <- bds1 %>%
  mutate(StrainID = fct_relevel(StrainID, 
                                "P1", "P2", "E1","E2","R",
                                "P1-E1","P1-E1-R","P1-E2","P1-E2-R","P1-R",
                                "P2-E1", "P2-E1-R","P2-E2","P2-E2-R","P2-R"))
```

### Fig 2A Part 1
```{r Figure 2A part 1, include = TRUE}
earlyestab1 <- bds1 %>% 
  #dplyr::filter(Grouping == "1") %>%
  group_by(Lineage_type, StrainID) %>%
  ggplot(aes(x = StrainID, y = bd1, color = as.factor(StrainID), fill = as.factor(StrainID), group = StrainID)) +
  geom_hline(yintercept =  as.numeric(bds1P1), linetype = "dashed", color = "#009E73", size = 0.5) +
  geom_hline(yintercept = as.numeric(bds1P2), linetype = "dotted", color = "#785EF0", size = 0.5) +
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +
  geom_errorbar(aes(ymin = bd1 - bd1se, ymax = bd1 + bd1se), width = 0,  size = 0.10) +
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +
  geom_line(alpha = 0.45) +
  ## Same problem as previous plotting chunks
  #scale_color_manual(values = col3, name = "StrainID")+
  #scale_fill_manual(values = col3a, name = "StrainID") +
  scale_shape_manual(values = shape3)+
  #singlefigtheme+
  ylim(-0.10, 175) +
  #facet_grid(~ Grouping, scales = 'free', labeller = labeller(Grouping = genolabel)) +
  theme_bw() +
  theme(legend.position = 'none', panel.grid.major = element_blank(), 
        panel.grid.minor =  element_blank()) +
  xlab("") + ylab("Competiveness (early Bd load)")+
  theme(axis.title = element_text(size = 10))+
  theme(text = element_text(size = 10))+
  #theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust = 0.5))+
  theme(axis.text.x = element_blank())+
  theme(axis.ticks = element_line(size = 0.10) ,
        axis.ticks.length = unit(0.10, "cm"))
earlyestab1
```

### Fig 2A Part 2
```{r Figure 2A part 2, include = TRUE}
earlyestab2 <- bds1 %>% 
  #dplyr::filter(PlotGroup == "1") %>%
  group_by(Lineage_type, StrainID) %>%
  ggplot(aes(x = StrainID, y = bd1, color = as.factor(StrainID), 
             fill = as.factor(StrainID), group = StrainID)) +
  geom_hline(yintercept =  as.numeric(bds1P1), linetype = "dashed", 
             color = "#009E73", size = 0.5) +
  geom_point(size = 2.50, aes(shape = StrainID, color = StrainID)) +
  geom_errorbar(aes(ymin = bd1 - bd1se, ymax = bd1 + bd1se), width = 0,  size = 0.10) +
  geom_rect(xmin = 0.85, xmax = 1.15, ymin = -Inf, ymax = Inf, 
            fill = "darkgrey", color = "NA", alpha = 0.05)+
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +
  geom_line(alpha = 0.45) +
  ## Same problem as previous plotting chunks
  #scale_color_manual(values = colP1, name = "StrainID")+
  #scale_fill_manual(values = colP1a, name = "StrainID") +
  #scale_shape_manual(values = shapeP1)+
  scale_shape_manual(values = shape3)+
  #singlefigtheme+
  ylim(-0.10, 175) +
  #facet_grid(~ Grouping, scales = 'free', labeller = labeller(Grouping = genolabel)) +
  theme_bw() +
  theme(legend.position = 'none', panel.grid.major = element_blank(), 
        panel.grid.minor =  element_blank()) +
  xlab("") + ylab("")+
  theme(axis.title = element_text(size = 10))+
  theme(text = element_text(size = 10))+
  theme(axis.text.y = element_blank())+
  #theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust = 0.5))+
  theme(axis.text.x = element_blank())+
  theme(axis.ticks = element_line(size = 0.10) ,
        axis.ticks.length = unit(0.10, "cm")) 
earlyestab2
```


### Fig 2A Part 3
```{r Figure 2A part 3, include = TRUE}


earlyestab3 <- bds1 %>% 
  #dplyr::filter(PlotGroup == "2") %>%
  group_by(Lineage_type, StrainID) %>%
  ggplot(aes(x = StrainID, y = bd1, color = as.factor(StrainID), 
             fill = as.factor(StrainID), group = StrainID)) +
  geom_hline(yintercept = as.numeric(bds1P2), linetype = "dotted", 
             color = "#785EF0", size = 0.5) +
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +
  geom_errorbar(aes(ymin = bd1 - bd1se, ymax = bd1 + bd1se), width = 0,  size = 0.10) +
  geom_rect(xmin = 0.85, xmax = 1.15, ymin = -Inf, ymax = Inf, 
            fill = "darkgrey", color = "NA", alpha = 0.05)+
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +
  geom_line(alpha = 0.45) +
  ## Same problem as previous plotting chunks
  #scale_color_manual(values = colP2, name = "StrainID")+
  #scale_fill_manual(values = colP2a, name = "StrainID") +
  #scale_shape_manual(values = shapeP2)+
  scale_shape_manual(values = shape3)+
  #singlefigtheme+
  ylim(-0.10, 175) +
  #facet_grid(~ Grouping, scales = 'free', labeller = labeller(Grouping = genolabel)) +
  theme_bw() +
  theme(legend.position = 'none', panel.grid.major = element_blank(), 
        panel.grid.minor =  element_blank()) +
  xlab("") + ylab("")+
  theme(axis.title = element_text(size = 10))+
  theme(text = element_text(size = 10))+
  theme(axis.text.y = element_blank())+
  #theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust = 0.5))+
  theme(axis.text.x = element_blank())+
  theme(axis.ticks = element_line(size = 0.10) ,
        axis.ticks.length = unit(0.10, "cm")) 
earlyestab3

 
```

### Final Fig 2A
```{r Combine figures for Fig 2A, include = TRUE}

Eestabl <- earlyestab1  + earlyestab2 + earlyestab3 +
  plot_layout(ncol = 3)


# BdPrev + plot_annotation(tag_levels = c('A', '1'), tag_prefix = '', tag_sep = '', 
#          tag_suffix = '') & 
#   theme(plot.tag.position = c(0.12, 0.98),
#         plot.tag = element_text(size = 12, hjust = 0, vjust = 0))
#   
ggsave("Eestabl.png", height = 3.5, width = 10, dpi = 600)

plot1 <- tiff("Eestabl.tiff", width = 15, height = 22, units = "cm", res = 300, pointsize = 10) 
plot1
dev.off() 

Eestabl + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
```

## Life Span Figure

```{r lifespan figure initialization, include = FALSE}

lifespan <- df %>%
  group_by(ID, StrainID, Lineage_type, Lineage) %>%
  filter(FieldN_2 == "S2") %>% 
  mutate(dod2  = sum(Day_swab, na.rm = T)) %>%
  select(ID, StrainID, Lineage_type, Lineage, dod2) %>%
  distinct() %>% as.data.frame()
lifespan



lifespanse <- lifespan %>% 
  group_by(StrainID, Lineage_type) %>%
  mutate(meandod = mean(dod2, na.rm = TRUE)) %>% 
  mutate(nobs = n()) %>% 
  mutate(dodse = sd(dod2, na.rm = TRUE)/sqrt(nobs)) %>%
  select(StrainID, Lineage_type, meandod, dodse, nobs) %>%
  distinct() %>% as.data.frame()
lifespanse


lifespanP1 <- as.list(subset(lifespanse, StrainID == "P1", meandod ))

lifespanP2 <- as.list(subset(lifespanse, StrainID == "P2", meandod))


# now, plot virulence on life span
# Re-order Lineage type for the correct figure order
lifespanse$Lineage_type <- factor(lifespanse$Lineage_type, c("Enzootic","Panzootic","Recombinant","E-P Coinfection","R Coinfection"))

lifespanse$StrainID <- factor(lifespanse$StrainID, c("P1", "P2", "E1", "E2", "R"))


genolabel <- c("Single Genotype Infections", "Enzootic-Panzootic Coinfections", "Recombinant-E or -P Coinfections", "Tripple Genotype (E-P-R) Coinfections") 
names(genolabel) <- c("1", "2", "3", "4")


lifespanse <- lifespanse %>%
  mutate(StrainID = fct_relevel(StrainID, 
                                "P1", "P2", "E1","E2","R",
                                "P1-E1","P1-E1-R","P1-E2","P1-E2-R","P1-R",
                                "P2-E1", "P2-E1-R","P2-E2","P2-E2-R","P2-R"))

```

### Lifespan Fig Part 1
```{r part 1, include = TRUE}



pls <- lifespanse %>% 
  #dplyr::filter(Grouping == "1") %>% 
  group_by(Lineage_type, StrainID) %>%
  ggplot(aes(x = StrainID, y = meandod, color = as.factor(StrainID), 
             fill = as.factor(StrainID), group = StrainID)) +
  geom_hline(yintercept =  as.numeric(lifespanP1), linetype = "dashed", 
             color = "#009E73", size = 0.5) +
  geom_hline(yintercept = as.numeric(lifespanP2), linetype = "dotted", 
             color = "#785EF0", size = 0.5) +
  geom_point(size = 2.50, aes(shape = StrainID, color = StrainID)) +
  geom_errorbar(aes(ymin = meandod - dodse, ymax = meandod + dodse), width = 0,  size = 0.10) +
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +
  geom_line(alpha = 0.45) +
  scale_color_manual(values = col3, name = "StrainID")+
  scale_fill_manual(values = col3a, name = "StrainID") +
  scale_shape_manual(values = shape3) +
  ylim(30, 75)+
  theme_bw() +
  theme(legend.position = 'none', panel.grid.major = element_blank(), 
        panel.grid.minor =  element_blank()) +
  xlab("") + ylab("Virulence (host life span, days post innoculation)")+
  theme(axis.title = element_text(size = 10))+
  theme(text = element_text(size = 10))+
  theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust = 0.5))+
  #theme(axis.text.x =  element_blank()) +
  theme(axis.ticks = element_line(size = 0.10) , 
        axis.ticks.length = unit(0.10, "cm")) 
pls


```

### Lifespan Fig Part 2
```{r part 2, include = TRUE}

pls2 <- lifespanse %>% 
  #dplyr::filter(PlotGroup == "1") %>% 
  group_by(Lineage_type, StrainID) %>%
  ggplot(aes(x = StrainID, y = meandod, color = as.factor(StrainID), 
             fill = as.factor(StrainID), group = StrainID)) +
  geom_hline(yintercept =  as.numeric(lifespanP1), linetype = "dashed", 
             color = "#009E73", size = 0.5) +
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +
  geom_errorbar(aes(ymin = meandod - dodse, ymax = meandod + dodse), width = 0,  size = 0.10) +
  geom_rect(xmin = 0.85, xmax = 1.15, ymin = -Inf, ymax = Inf, 
            fill = "darkgrey", color = "NA", alpha = 0.05)+
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +
  geom_line(alpha = 0.45) +
  scale_color_manual(values = colP1, name = "StrainID")+
  scale_fill_manual(values = colP1a, name = "StrainID") +
  scale_shape_manual(values = shapeP1) +
  ylim(30, 75)+
  theme_bw() +
  theme(legend.position = 'none', panel.grid.major = element_blank(), 
        panel.grid.minor =  element_blank()) +
  xlab("") + ylab("")+
  theme(axis.title = element_text(size = 10))+
  theme(text = element_text(size = 10))+
  theme(axis.text.y = element_blank())+
  theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust = 0.5))+
  #theme(axis.text.x =  element_blank()) +
  theme(axis.ticks = element_line(size = 0.10) , 
        axis.ticks.length = unit(0.10, "cm")) 
pls2

```

### Lifespan Fig Part 3
```{r part 3, include = TRUE}

pls3 <- lifespanse %>% 
  #dplyr::filter(PlotGroup == "2") %>% 
  group_by(Lineage_type, StrainID) %>%
  ggplot(aes(x = StrainID, y = meandod, color = as.factor(StrainID), 
             fill = as.factor(StrainID),  group = StrainID)) +
  geom_hline(yintercept = as.numeric(lifespanP2), linetype = "dotted", 
             color = "#785EF0", size = 0.5) +
  geom_point(size = 2.50, aes(shape = StrainID, color = StrainID)) +
  geom_errorbar(aes(ymin = meandod - dodse, ymax = meandod + dodse), width = 0,  size = 0.10) +
  geom_rect(xmin = 0.85, xmax = 1.15, ymin = -Inf, ymax = Inf, 
            fill = "darkgrey", color = "NA", alpha = 0.05)+
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +
  geom_line(alpha = 0.45) +
  scale_color_manual(values = colP2, name = "StrainID")+
  scale_fill_manual(values = colP2a, name = "StrainID") +
  scale_shape_manual(values = shapeP2) +
  #singlefigtheme+
  ylim(30, 75)+
  #facet_grid(~ Grouping, scales = 'free', labeller = labeller(Grouping = genolabel)) +
  theme_bw() +
  theme(legend.position = 'none', panel.grid.major = element_blank(), 
        panel.grid.minor =  element_blank()) +
  xlab("") + ylab("")+
  theme(axis.title = element_text(size = 10))+
  theme(text = element_text(size = 10))+
  theme(axis.text.y = element_blank())+
  #theme(axis.text.x =  element_blank()) +
  theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust = 0.5))+
  theme(axis.ticks = element_line(size = 0.10) , 
        axis.ticks.length = unit(0.10, "cm")) 
pls3

```


### Lifespan Fig Final
```{r, include = TRUE}
Vir <- pls  + pls2 + pls3 +
  plot_layout(ncol = 3)
Vir

# BdPrev + plot_annotation(tag_levels = c('A', '1'), tag_prefix = '', tag_sep = '', 
#          tag_suffix = '') & 
#   theme(plot.tag.position = c(0.12, 0.98),
#         plot.tag = element_text(size = 12, hjust = 0, vjust = 0))
#   
ggsave("Vir.png", height = 3.5, width = 10, dpi = 600)

plot1 <- tiff("Vir.tiff", width = 15, height = 22, units = "cm", res = 300, pointsize = 10) 
plot1
dev.off() 
```


## Figure 3  - Competitiveness (relative BdGPL1) with virulence on the x-axis
```{r FIGURE 3 setup, include = FALSE}


relbdp1 <- df %>%
  group_by(ID, StrainID, Lineage_type, Lineage, Treatment, FieldN_2,  Day_swab, TotalBd) %>%
  dplyr::filter(Treatment == "Coinfected") %>% 
  mutate(meanGPL = mean(GPL_pg, na.rm = T)) %>%
  mutate(totGPL  = sum(GPL_pg, na.rm = T)) %>%
  mutate(meanBZL = mean(Brazil_pg, na.rm = T)) %>%
  mutate(totBZL  = sum(Brazil_pg, na.rm = T)) %>%
  mutate(meanH   = mean(Hybrid , na.rm = T)) %>%
  mutate(totH    = sum(Hybrid , na.rm = T)) %>%
  select(ID, StrainID, Lineage_type, Lineage, Treatment, FieldN_2,  Day_swab, meanGPL, totGPL, meanBZL, totBZL, meanH, totH, TotalBd) %>%
  distinct() %>% as.data.frame()
head(relbdp1)



relbdp2 <- relbdp1 %>% 
  group_by(ID, StrainID, Lineage_type, Lineage, Treatment, FieldN_2, Day_swab) %>% 
  mutate(totbd   = sum(totGPL + totBZL + totH, na.rm = T)) %>%
  mutate(relgpl  = (totGPL/TotalBd)) %>%
  group_by(StrainID) %>%
  mutate(n_obs = n()) %>% # count the number of observations per group
  select(ID, StrainID, Lineage_type, Lineage, Treatment, FieldN_2, Day_swab, relgpl) %>%
  distinct() %>% as.data.frame()
head(relbdp2)


relbdp3 <- relbdp2 %>% 
  group_by(StrainID, Treatment, FieldN_2) %>%
  filter(Treatment == "Coinfected") %>% 
  filter(FieldN_2 == "S2") %>% 
  mutate(dayofdeath = mean(Day_swab, na.rm = TRUE)) %>% 
  mutate(meanrelgpl = mean(relgpl, na.rm = TRUE)) %>% 
  mutate(nobs = n()) %>% 
  mutate(relgplse = sd(relgpl, na.rm = TRUE)/sqrt(nobs)) %>%
  mutate(meandod = mean(Day_swab, na.rm = TRUE)) %>% 
  mutate(nobs = n()) %>% 
  mutate(dodse = sd(Day_swab, na.rm = TRUE)/sqrt(nobs)) %>%
  select(StrainID, Lineage_type, meandod, dodse, meanrelgpl, relgplse, dayofdeath) %>%
  distinct() %>% as.data.frame()
relbdp3


# Re-order Lineage type for the correct figure order
genolabel <- c("Single Genotype Infections", "Panzootic-Enzootic Coinfections", "Recombinant-E or -P Coinfections", "Tripple Genotype (P-E-R) Coinfections") 
names(genolabel) <- c("1", "2", "3", "4")


# now plot competitiveness 
pcomp1 <- relbdp3 %>% 
  group_by(Lineage_type, StrainID) %>% 
  ggplot(aes(x = meandod, y = meanrelgpl, group =  StrainID)) +
  geom_point(size = 3.50)  +
  geom_errorbar(aes(ymin = meanrelgpl - relgplse, ymax = meanrelgpl + relgplse),  width = 0,   size = 0.10) +
  geom_line(alpha = 0.45) +
  xlim(30, 70) +
  #singlefigtheme+
  geom_point(stat = "identity", position = position_dodge(), inherit.aes = TRUE, alpha = 0.5)+
  geom_errorbarh(aes(xmin = meandod - dodse, xmax = meandod + dodse), position = position_dodge(.9), width = 0, height = 0, size = 0.10, na.rm = T)+
  #facet_wrap(~ Lineage_type, scales = 'free') +
  #facet_grid(~ Grouping, scales = 'free', labeller = labeller(Grouping = genolabel))+
  theme_bw() +
  theme(legend.position = 'none', panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("Virulence (host life span, days post innoculation)") + ylab("Competitive outcome (relative P1 or P2)")+
  theme(axis.title = element_text(size = 10))+
  theme(text = element_text(size = 10))+
  theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust = 1))+
  theme(axis.ticks = element_line(size = 0.10) , 
        axis.ticks.length = unit(0.10, "cm")) +
  theme(strip.background = element_rect(fill = "white", color = "black", size = 0.25),
        strip.text = element_text(colour = "black", size = 8),
        panel.border =  element_rect(color = "black", fill = NA, size = 0.25))

ggsave("pcomp1.png", pcomp1, height = 3.5, width = 10, dpi = 600)
ggsave("pcomp1.eps", pcomp1, height = 3.55, width = 10, dpi = 600)
```

```{r FIGURE 3, include = TRUE}
pcomp1 + 
 geom_label(aes(label = StrainID))
# 





BdPlotComp <- plot_grid(earlyestab1, earlyestab2, earlyestab3, nrow = 1, align = "hv")
BdPlotComp 


#BdPlot1b <- plot_grid(RTPlot1, BdPlotComp, BdPlot1a, nrow = 3, align = "l")
#BdPlot1b



#### COMBINE PLOTS #### 

p1 <- plot_grid(
  earlyestab1, earlyestab2, earlyestab3,
  nrow = 1,
  label_size = 10,
  align = "l",
  axis = "tb")
p1

p2 <- plot_grid(
  pls, pls2, pls3,
  nrow = 1,
  label_size = 10,
  align = "l",
  axis = "tb")
p2


p3 <- plot_grid(
  p1, p2, 
  nrow = 2,
  label_size = 10,
  align = "l",
  axis = "tb")
p3



p3 <- plot_grid(
  p3, pcomp1,
  nrow = 1,
  label_size = 10,
  align = "l",
  axis = "tb")
p3

#ggsave("BdplotA.png", BdPlot1b, height = 10, width = 12, dpi = 600, family = "Arial")

```


```{r candidate models, include = FALSE}
# lm1<- glm(log(meandod) ~ as.factor(StrainID), data = P1life)
# lm1
# summary.lm(lm1)
# summary(lm1)
# plot(lm1)

```

## Figure 4 - TOTAL Bd Load (per host)
```{r FIGURE 4 setup, include = FALSE}

#### Figure 4 - TOTAL Bd Load (per host) ####



#### MEAN Bd Per Host #### 
summary_bd1 <- df %>%
  group_by(ID, StrainID, Lineage_type, Lineage, Mortality_percentage, Cleared_percentage) %>%
  mutate(meanGPL = mean(GPL_pg, na.rm = T)) %>%
  mutate(totGPL  = sum(GPL_pg, na.rm = T)) %>%
  mutate(meanBZL = mean(Brazil_pg, na.rm = T)) %>%
  mutate(totBZL  = sum(Brazil_pg, na.rm = T)) %>%
  mutate(meanHybrid = mean(Hybrid, na.rm = T)) %>%
  mutate(totHyb  = sum(Hybrid, na.rm = T)) %>%
  select(ID, StrainID, Lineage_type, Lineage, Mortality_percentage, Cleared_percentage, meanGPL, totGPL, meanBZL, totBZL, meanHybrid, totHyb) %>%
  distinct() %>% as.data.frame()
head(summary_bd1)


summary_bd2 <- summary_bd1 %>% 
  group_by(ID, StrainID, Lineage_type, Lineage, Mortality_percentage, Cleared_percentage) %>% 
  mutate(totbd   = sum(totGPL + totBZL + totHyb, na.rm = T)) %>%
  mutate(relGPL  = (totGPL/totbd)) %>%
  group_by(StrainID) %>%
  dplyr::mutate(n_obs = n()) %>% # count the number of observations per group
  mutate(bdse = sd(totbd, na.rm = TRUE)/sqrt(n())) %>%
  select(ID, StrainID, Lineage_type, Lineage, Mortality_percentage, Cleared_percentage, meanGPL, totGPL, meanBZL, totBZL, meanHybrid, totHyb, totbd, bdse, relGPL) %>%
  distinct() %>% as.data.frame()
head(summary_bd2)




#### MEANS AND SE FOR PLOTTING ####
summary_se <- summary_bd2 %>% 
  group_by(StrainID, Lineage_type, ) %>%
  mutate(n_obs = n()) %>% # count the number of observations per group
  mutate(meanbd = mean(totbd, na.rm = TRUE)) %>% 
  mutate(bdse = sd(totbd, na.rm = TRUE)/sqrt(n())) %>%
  select(StrainID, Lineage_type, meanbd, bdse) %>%
  distinct() %>% as.data.frame()
summary_se

summary_se <- summary_se %>%
  mutate(StrainID = fct_relevel(StrainID, 
                                "P1", "P2", "E1","E2","R",
                                "P1-E1","P1-E1-R","P1-E2","P1-E2-R","P1-R",
                                "P2-E1", "P2-E1-R","P2-E2","P2-E2-R","P2-R"))



p1meanbd <- as.list(subset(summary_se, StrainID == "P1", meanbd))
p1meanbd

p2meanbd <- as.list(subset(summary_se, StrainID == "P2", meanbd))
p2meanbd


summary_bd3 <- summary_bd2 %>% 
  group_by(StrainID, Mortality_percentage) %>% 
  mutate(totbd   = sum(totGPL + totBZL, na.rm = T)) %>%
  mutate(relGPL  = (totGPL/totbd)) %>%
  select(StrainID, Mortality_percentage, Cleared_percentage, totbd) %>%
  distinct() %>% as.data.frame()
summary_bd3





# Re-order Lineage type for the correct figure order
summary_se$Lineage_type <- factor(summary_se$Lineage_type, c("Enzootic","Panzootic","Recombinant","E-P Coinfection","R Coinfection"))
genolabel <- c("Single Genotype Infections", "Enzootic-Panzootic Coinfections", "Recombinant-E or -P Coinfections", "Tripple Genotype (E-P-R) Coinfections") 
names(genolabel) <- c("1", "2", "3", "4")


#png(file = "TotalBdLoad.png", height = 3, width = 3.5, units = 'in', res = 300)
```

```{r FIGURE4}
plotmeanbd1 <- summary_se %>% 
  #dplyr::filter(Grouping == "1") %>%
  group_by(Lineage_type, StrainID) %>%
  ggplot(aes(x = StrainID, y = meanbd, color = as.factor(StrainID), fill = as.factor(StrainID), group = StrainID)) +
  geom_hline(yintercept =  as.numeric(p1meanbd), linetype = "dashed", color = "#009E73", size = 0.5) +
  geom_hline(yintercept = as.numeric(p2meanbd), linetype = "dotted", color = "#785EF0", size = 0.5) +
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +  
  geom_errorbar(aes(ymin = meanbd - bdse, ymax = meanbd + bdse), width = 0,  size = 0.10) +
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +  
  geom_line(alpha = 0.45) +
  ## Same problem as previous plotting chunks
  #scale_color_manual(values = col3, name = "StrainID")+
  #scale_fill_manual(values = col3a, name = "StrainID") +
  scale_shape_manual(values = shape3) +
  #singlefigtheme+
  ylim(20, 800) +
  #facet_grid(~ Grouping, scales = 'free', labeller = labeller(Grouping = genolabel)) +
  theme_bw() +
  theme(legend.position = 'none', panel.grid.major = element_blank(), 
        panel.grid.minor =  element_blank()) +
  xlab("") + ylab("Total Bd load (per host)")+
  theme(axis.title = element_text(size = 10))+
  theme(text = element_text(size = 10))+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5))+
  theme(axis.ticks = element_line(size = 0.10) , 
        axis.ticks.length = unit(0.10, "cm")) +
  theme(strip.background = element_rect(fill = "white", color = "black", size = 0.25),
        strip.text.x = element_blank(),
        panel.border =  element_rect(color = "black", fill = NA, size = 0.25))
plotmeanbd1



plotmeanbd2 <- summary_se %>% 
  #dplyr::filter(PlotGroup == "1") %>%
  group_by(Lineage_type, StrainID) %>%
  ggplot(aes(x = StrainID, y = meanbd, color = as.factor(StrainID), fill = as.factor(StrainID), group = StrainID)) +
  geom_hline(yintercept =  as.numeric(p1meanbd), linetype = "dashed", color = "#009E73", size = 0.5) +
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +  
  geom_errorbar(aes(ymin = meanbd - bdse, ymax = meanbd + bdse), width = 0,  size = 0.10) +
  geom_rect(xmin = 0.85, xmax = 1.15, ymin = -Inf, ymax = Inf, 
            fill = "darkgrey", color = "NA", alpha = 0.05)+
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +  
  geom_line(alpha = 0.45) +
  ## Same problem as previous plotting chunks
  #scale_color_manual(values = colP1, name = "StrainID")+
  #scale_fill_manual(values = colP1a, name = "StrainID") +
  #scale_shape_manual(values = shapeP1) +
  scale_shape_manual(values = shape3)+
  #singlefigtheme+
  ylim(20, 800) +
  #facet_grid(~ Grouping, scales = 'free', labeller = labeller(Grouping = genolabel)) +
  theme_bw() +
  theme(legend.position = 'none', panel.grid.major = element_blank(), 
  panel.grid.minor =  element_blank()) +
  xlab("") + ylab("")+
  theme(axis.title = element_text(size = 10))+
  theme(text = element_text(size = 10))+
  theme(axis.text.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5))+
  theme(axis.ticks = element_line(size = 0.10) , 
        axis.ticks.length = unit(0.10, "cm")) +
  theme(strip.background = element_rect(fill = "white", color = "black", size = 0.25),
        strip.text.x = element_blank(),
        panel.border =  element_rect(color = "black", fill = NA, size = 0.25))
plotmeanbd2




plotmeanbd3 <- summary_se %>% 
  #dplyr::filter(PlotGroup == "2") %>%
  group_by(Lineage_type, StrainID) %>%
  ggplot(aes(x = StrainID, y = meanbd, color = as.factor(StrainID), fill = as.factor(StrainID), group = StrainID)) +
  geom_hline(yintercept = as.numeric(p2meanbd), linetype = "dotted", color = "#785EF0", size = 0.5) +
  geom_point(size = 2.50, aes(shape = StrainID, color = StrainID)) +  
  geom_errorbar(aes(ymin = meanbd - bdse, ymax = meanbd + bdse), width = 0,  size = 0.10) +
  geom_rect(xmin = 0.85, xmax = 1.15, ymin = -Inf, ymax = Inf, 
            fill = "darkgrey", color = "NA", alpha = 0.05)+
  geom_point(size = 3.50, aes(shape = StrainID, color = StrainID)) +  
  geom_line(alpha = 0.45) +
  ## Same problem as previous plotting chunks
  #scale_color_manual(values = colP2, name = "StrainID")+
  #scale_fill_manual(values = colP2a, name = "StrainID") +
  #scale_shape_manual(values = shapeP2) +
  scale_shape_manual(values = shape3)+
  #singlefigtheme+
  ylim(20, 800) +
  #facet_grid(~ Grouping, scales = 'free', labeller = labeller(Grouping = genolabel)) +
  theme_bw() +
  theme(legend.position = 'none', panel.grid.major = element_blank(), 
        panel.grid.minor =  element_blank()) +
  xlab("") + ylab("")+
  theme(axis.title = element_text(size = 10))+
  theme(text = element_text(size = 10))+
  theme(axis.text.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5))+
  theme(axis.ticks = element_line(size = 0.10) , 
        axis.ticks.length = unit(0.10, "cm")) +
  theme(strip.background = element_rect(fill = "white", color = "black", size = 0.25),
        strip.text.x = element_blank(),
        panel.border =  element_rect(color = "black", fill = NA, size = 0.25))
plotmeanbd3


#### End Mean Bd Per Host #### 


#### END FIGURE 2b ####


#### COMBINE FIGURES FOR FIGURE TRANSMISSION ####

Trans <- plotmeanbd1  + plotmeanbd2 + plotmeanbd3 +
  plot_layout(ncol = 3)
Trans

# BdPrev + plot_annotation(tag_levels = c('A', '1'), tag_prefix = '', tag_sep = '', 
#          tag_suffix = '') & 
#   theme(plot.tag.position = c(0.12, 0.98),
#         plot.tag = element_text(size = 12, hjust = 0, vjust = 0))
#   
ggsave("Trans.png", height = 3.5, width = 10, dpi = 600)
```




